---
title: "Gov 50 Final Project"
author: "Lillian Cooper"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Introduction

I am interested in exploring data related to... how abortions and miscarriages are related in California 

Can the abortion rate predict the miscarriage rate? Abortions often involve intense medical procedures that can negatively affect a woman for future pregnancies. In addition, abortions are often found to be the source of psychological trauma for the mother(Pourreza and Aziz). This psychological trauma can lead to something called allostatic load, which is essentially the mechanism for how extreme stress for long periods of time corrodes health (Roder). Due to potential physical damages of an abortion and the psychological damages caused by abortions, I hypothesize that the California abortion rate may be able to predict the miscarriage rate. 
I will be looking at a data set which describes the abortion rate and the number of miscarriages for women in California age 20-24 from 1988-2017. The abortion rate is measured per 1,000 women and my explanatory variable is the abortion rate.
If I observe a strong positive coefficient between abortions and miscarriages, then I know my hypothesis will be supported. (MAY GO DEEPER W THIS, maybe write what pattern would disprove the hypothesis) 

## Loading Data

```{r Loading and Cleaning Data}
library(tidyverse)
library(readr)

NationalAndStatePregnancy_PublicUse <- read_csv("NationalAndStatePregnancy_PublicUse.csv")


head(NationalAndStatePregnancy_PublicUse)
abortion_and_mis <- NationalAndStatePregnancy_PublicUse |>
  select(state, year, abortionratelt20: abortionrate40plus, miscarriageslt20: miscarriages40plus)


CA_abortion_and_mis <- abortion_and_mis |>
  subset(state == "CA")
CA_abortion_and_mis
```
## Abortion Rate in California over time

Based on the trend line, the abortion rate among women in California aged 20-24 has steadily decreased since the 90s.

```{r Abortion Rate Over Time}
library(dplyr)
library(broom)


CAabort_plot <- CA_abortion_and_mis |> 
  ggplot(mapping = aes(y = abortionrate2024)) + geom_point(mapping = aes(x = year)) + labs(x = "Year", y = "Abortion Rate Among Women in CA Age 20-24 (per 1,000)") +  geom_smooth(mapping = aes(x = year), method = "lm")
CAabort_plot

```

## Miscarriages in Califonia Overtime

This plot also reveals that miscarriages have decreased among California women aged 20-24 since the 90s.

```{r Miscarriage Count Over Time}
CAmis_plot <- CA_abortion_and_mis |>
  ggplot(mapping = aes(y = miscarriages2024)) + geom_point(mapping = aes(x = year)) + labs(x = "Year", y = "Miscarriages Among Women in CA age 20-24") + geom_smooth(mapping = aes(x = year), method = "lm")
CAmis_plot

```

## Miscarriages vs Abortion Rate

```{r Miscarriages vs Abortion Rate}
CAtogether_plot <- CA_abortion_and_mis |>
  ggplot(mapping = aes(y = miscarriages2024)) + geom_point(mapping = aes(x = abortionrate2024)) + labs(x = "Abortion Rate Among Women in CA Age 20-24 (per 1,000)", y = "Miscarriages Among Women in CA age 20-24") +  geom_smooth(mapping = aes(x = abortionrate2024), method = "lm")
CAtogether_plot

```

## Regression Analysis

The intercept on the Miscarriages vs Abortion Rate graph is approximately 13,046. This means that if the abortion rate in California among women aged 20-24 was 0, then we would expect the number of miscarriages to be about 13,046. The Abortion rate coefficient is 308 which means that, on average, my model predicts that for every 1 abortion / 1,000 women increase, the number of miscarriages would increase by 308. 

```{r Linear Regression}
regression <- lm(miscarriages2024 ~ abortionrate2024, data = CA_abortion_and_mis)
regression

varnames <- c("(Intercept)" = "Intercept", "abortionrate2024" = "Abortion Rate")


modelsummary :: modelsummary(regression, gof_map = c("nobs", "r.squared"), coef_map = varnames)

```

## Conclusion

Based on the R^2 value being 0.922, I see that there is a very strong positive correlation between abortion rates and number of miscarriages. This supports my hypothesis that the California abortion rate can predict the miscarriage rate. However, this is an observational study over time, and it is very important to remember that correlation does not mean causation. 
Some limitations of this data are that this only looks at women aged 20-24 which is a limited range. Expanding the range could greatly affect the trend line. Also, this is only in California. Different states and localities could show drastically different trends.
Another limitation of this data is that the same number of women who got abortions are not guarunteed to be the same women who has miscarriages. 
If I had more time and money, I would create a survey of women who had abortions and then follow up to see if thye had a miscarriage. However, this could be a little unethical asking women extremely personal questions. 
Overall, there is a strong correlation between abortion rate and miscarriages among women in California ages 20-24 from 1988-2017.



## Work Cited

Pourreza, Abolghasem, and Aziz Batebi. “Psychological Consequences of Abortion among the Post Abortion Care Seeking Women in Tehran.” Iranian journal of psychiatry vol. 6,1 (2011): 31-6.

Roder, Amy. “America Is Failing Its Black Mothers.” Harvard Public Health Magazine, 10 June 2020, www.hsph.harvard.edu/magazine/magazine_article/america-is-failing-its-black-mothers/. 




