---
title: "Gov 50 Final Project"
author: "Lillian Cooper"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Introduction

I am interested in exploring data related to... how abortions and miscarriages are related in California 

Can the abortion rate predict the miscarriage rate? Abortions often involve intense medical procedures that can negatively affect a woman for future pregnancies. In addition, abortions are often found to be the source of psychological trauma for the mother(CITE). This psychological trauma can lead to something called allostatic load, which is essentially the mechanism for how extreme stress for long periods of time corrodes health (Roder). Due to potential physical damages of an abortion and the psychological damages caused by abortions, I hypothesize that the California abortion rate may be able to predict the miscarriage rate. 
I will be looking at a data set which describes the abortion rate and the number of miscarriages for women in California age 20-24 from 1988-2017. The abortion rate is measured per 1,000 women and my explanatory variable is the abortion rate.
If I observe a strong positive coefficient between abortions and miscarriages, then I know my hypothesis will be supported. (MAY GO DEEPER W THIS, maybe write what pattern would disprove the hypothesis) 


```{r Loading and Cleaning Data}
library(tidyverse)
library(readr)
head(NationalAndStatePregnancy_PublicUse)

abortion_and_mis <- NationalAndStatePregnancy_PublicUse |>
  select(state, year, abortionratelt20: abortionrate40plus, miscarriageslt20: miscarriages40plus)
head(abortion_and_births)


CA_abortion_and_mis <- abortion_and_mis |>
  subset(state == "CA")
CA_abortion_and_mis
```
## Abortion Rate in CA over time

```{r Abortion Rate Over Time}
library(dplyr)
library(broom)


CAabort_plot <- CA_abortion_and_mis |> 
  ggplot(mapping = aes(y = abortionrate2024)) + geom_point(mapping = aes(x = year)) + labs(x = "Year", y = "Abortion Rate Among Women in CA Age 20-24 (per 1,000)") +  geom_smooth(mapping = aes(x = year), method = "lm")
CAabort_plot

```

## Miscarriage Plot

```{r Miscarriage Count Over Time}
CAmis_plot <- CA_abortion_and_mis |>
  ggplot(mapping = aes(y = miscarriages2024)) + geom_point(mapping = aes(x = year)) + labs(x = "Year", y = "Miscarriages Among Women in CA age 20-24") + geom_smooth(mapping = aes(x = year), method = "lm")
CAmis_plot

```

## Together plot

```{r Miscarriages vs Abortion Rate}
CAtogether_plot <- CA_abortion_and_mis |>
  ggplot(mapping = aes(y = miscarriages2024)) + geom_point(mapping = aes(x = abortionrate2024)) + labs(x = "Abortion Rate Among Women in CA Age 20-24 (per 1,000)", y = "Miscarriages Among Women in CA age 20-24") +  geom_smooth(mapping = aes(x = abortionrate2024), method = "lm")
CAtogether_plot

```

## Linear regression lol

```{r Linear Regression}
regression <- lm(miscarriages2024 ~ abortionrate2024, data = CA_abortion_and_mis)
regression

```

## How many pregnancies overall in California overtime? (may not be very useful)

```{r Pregnancies Overall}
CA_pregnancies <- NationalAndStatePregnancy_PublicUse |>
  select(state, year, pregnancyrate2024) |>
  subset(state == "CA")
head(CA_pregnancies)

CA_pregnancies |> 
  ggplot(mapping = aes(x = year)) + geom_point(mapping = aes(y = pregnancyrate2024))
  


```
