---
title: "Gov 50 Final Project"
author: "Lillian Cooper"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to... Sports statistics in major league baseball, finding a correlation mental health advocacy and sports performace, or religious identity and how it correlates to mental health amongst young adults 

Are lefty catchers in baseball really that bad? I other words, what is the success rate of lefty catchers in major league baseball compared to right catchers- and is this comparison statistically significant. 
I hypothesize that right-handed catchers will have a higher success rate in throwing out runners at second base than lefty catchers. 
My sample will be all starting major league right handed catchers from 1980-2023 and all starting lefty catchers from 1980-2023. I will then use a simple random sample of each set of catchers to narrow it down to be 10 righty catchers and 10 lefty catchers. The variabel will be =1 for righty and =0 for lefty.
The outcome variable will be the success rate a throwing the runner out at second base. I might also see if I can find data on framing success in baseball for righty and lefty catchers, but so far I don't exactly know the scale for my outcome variable.
If I observe similar success rates of throwing runners out from a righty and lefty catcher, then the common rumor that being a lefty catcher is bad will be disproved. If I observe the opposite, then the rumor will be true.

```{r}
if (!requireNamespace('pacman', quietly = TRUE)){
  install.packages('pacman')
}
pacman::p_load_current_gh("BillPetti/baseballr")
```

```{r}
library(tidyverse)
library(broom)
library(haven)
data <- bref_daily_batter("2015-08-01", "2015-10-03") 
data |>
  head()
```

```{r}
first_plot <- data |>
  ggplot(mapping = aes(x = BB, y = RBI)) + geom_point() + labs( x = "# of Walks Aug.- Oct. 2015", y = "# of RBIs Aug.- Oct. 2015")
first_plot
  

```